---
title: Single-Cell RNA Sequencing Distinguishes Heterogenous Transcriptional Signatures
  in Macrophages during Efferocytosis
author: "Connor Lantz, Behram Radmanesh, Esther Liu, Edward B. Thorp, Jennie Lin"
date: Northwestern University, Feinberg School of Medicine, Departments of Pathology and Medicine and the Feinberg Cardiovascular and Renal Research Institute, Jesse Brown Veterans Affairs Medical Center, Chicago IL, 60611
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r include = FALSE}
library(pkgbuild)
library(ggplot2)
library(dplyr)
library(Matrix)
library(sctransform)
library(cowplot)
library(monocle)
library(Seurat)
library(umap)
library(RCurl)
library(scales)
library(tidyverse)
library(MAST)
library(RColorBrewer)
library(destiny)
library(SingleCellExperiment)
library(ggbeeswarm)
library(ggthemes)
library(slingshot)
library(clusterExperiment)
library(gam)
library(scater)
library(gridExtra)
library(GOplot)
```

```{r include=FALSE}
setwd("C:/Users/conno/Documents/Driskill Graduate Program/Projects/Efferocytosis Containaminating RNA/scRNAseq/MANUSCRIPT_scRNAseq")
```
# Abstract
Efferocytosis triggers cellular reprogramming, including transcripts that encode anti-inflammatory cytokines that promote inflammation resolution. Our current understanding of this transcriptional response is largely informed from analysis of bulk phagocyte populations. This precludes the resolution of heterogeneity between individual macrophages and macrophage population subsets. Moreover, phagocytes may be contaminated by passenger transcripts that originate from engulfed apoptotic bodies. To define the transcriptional diversity of efferocytosis, we utilized single cell mRNA sequencing after co-cultivation with apoptotic cells. Importantly, transcriptomic analysis was performed after validating degradation of apoptotic cell-derived transcripts. Our findings reveal heterogeneity of the efferocytic response at single-cell resolution, particularly evident between F4/80+ MHCIILO and F4/80- MHCIIHI macrophage sub-populations. F4/80+ MHCIILO macrophages induced pathways associated with tissue homeostasis and response to oxidative stress, while the F4/80- MHCIIHI subset downregulated these pathways. Ablation of a canonical efferocytosis receptor, MerTK, blunted efferocytic signatures, and led to elevation of apoptotic pathways in F4/80+ MHCIILO macrophages after challenge with apoptotic cells, thereby offering a validation of our approach. Our results highlight the utility of single-cell transcriptomics to study heterogenous transcriptional responses in macrophage subpopulations after exposure to apoptotic cells. 

# Removing Low-quality Cells and Integrating Single Cell Datasets
## Create Seurat Objects

For each condition's dataset, a Seurat Object is created by reading the corresponding 10X dataset. Low quality cells with less than 100 features are excluded from each Seurat object.

```{r Create Seurat Objects}
for (file in c("WT_Ctrl", "WT_2_hr", "WT_6_hr", "MerKD_Ctrl", "MerKD_2hr", "MerKD_6_hr")){
  seurat_data <- Read10X(data.dir = paste0("Data/", file))
  seurat_obj <- CreateSeuratObject(counts = seurat_data, min.features = 100, project = file)
  assign(file, seurat_obj)
}
```

Each Seurat Object will now be merged to create a single Seurat Object which will be utlilized for quality control analyses.
```{r}
#Merge Seurat Object for QC
merged_seurat <- merge(x = WT_Ctrl, y = c(WT_2_hr, WT_6_hr, MerKD_Ctrl, MerKD_2hr, MerKD_6_hr), add.cell.ids = c("WT_Ctrl", "WT_2_hr", "WT_6_hr", "MerKD_Ctrl", "MerKD_2hr", "MerKD_6_hr"))

merged_seurat
```

## Quality Control 
```{r Create Metadata}
#Log Normalization of Merged Seurat
merged_seurat$log10GenesPerUMI <- log10(merged_seurat$nFeature_RNA) / log10(merged_seurat$nCount_RNA)

#Mitochondrial Feature Expression Ratio
merged_seurat$mitoRatio <- PercentageFeatureSet(merged_seurat, pattern = "^mt-")
merged_seurat$mitoRatio <- merged_seurat@meta.data$mitoRatio / 100

#Ribosomal Feature Expression Ratio
merged_seurat$riboRatio <- PercentageFeatureSet(merged_seurat, pattern = "^Rp[ls]")
merged_seurat$riboRatio <- merged_seurat@meta.data$riboRatio / 100

#Create Metadata Slot in Merged Seurat
metadata <- merged_seurat@meta.data
metadata$cells <- rownames(metadata)
metadata <- metadata %>%
  dplyr::rename(seq_folder = orig.ident,
                nUMI = nCount_RNA,
                nGene = nFeature_RNA)

#Add Identification Metadata for Each Condition
metadata$sample <- NA
metadata$sample[which(str_detect(metadata$cells, "^WT_Ctrl"))] <- "WT_Ctrl"
metadata$sample[which(str_detect(metadata$cells, "^WT_2_hr"))] <- "WT_2_hr"
metadata$sample[which(str_detect(metadata$cells, "^WT_6_hr"))] <- "WT_6_hr"
metadata$sample[which(str_detect(metadata$cells, "^MerKD_Ctrl"))] <- "MerKD_Ctrl"
metadata$sample[which(str_detect(metadata$cells, "^MerKD_2hr"))] <- "MerKD_2hr"
metadata$sample[which(str_detect(metadata$cells, "^MerKD_6_hr"))] <- "MerKD_6_hr"
merged_seurat@meta.data <- metadata
```

Different quality control metrics such as the number of UMIs, genes, mitochondrial, and ribosomal counts are analyzed to determine the quality of the data. The results from these different quality control metrics will be used to filter out low-quality cells.

```{r Quality Control Metrics, echo=TRUE, warning=FALSE}
#nCells
p1 <- metadata %>% 
  ggplot(aes(x=sample, fill=sample)) + 
  geom_bar() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("nCells")

#nUMIs
p2 <- metadata %>% 
  ggplot(aes(color=sample, x=nUMI, fill= sample)) + 
  geom_density(alpha = 0.2) + 
  scale_x_log10() + 
  theme_classic() +
  ylab("Cell density") +
  geom_vline(xintercept = 1000) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("nUMIS")

#nGenes 
p3 <- metadata %>% 
  ggplot(aes(x=sample, y=log10(nGene), fill=sample)) + 
  geom_boxplot() + 
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("nGenes")

#mito counts ratio
p4 <- metadata %>% 
  ggplot(aes(color=sample, x=mitoRatio, fill=sample)) + 
  geom_density(alpha = 0.1) + 
  scale_x_log10() + 
  theme_classic() +
  geom_vline(xintercept = 0.1) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("Mito Ratio")

#ribo counts ratio
p5 <- metadata %>% 
  ggplot(aes(color=sample, x=riboRatio, fill=sample)) + 
  geom_density(alpha = 0.2) + 
  scale_x_log10() + 
  theme_classic() +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("Ribo Ratio")

#novelty gene
p6 <- metadata %>%
  ggplot(aes(x=log10GenesPerUMI, color = sample, fill=sample)) +
  geom_density(alpha = 0.2) +
  theme_classic() +
  geom_vline(xintercept = 0.8) +
  theme(plot.title = element_text(hjust=0.5, face="bold")) +
  ggtitle("Unique genes per UMI")

grid.arrange(p1,p2,p3,p4,p5,p6, ncol = 2)
```

```{r Genes versus UMI Correlations, echo=TRUE, warning=FALSE}
#Correlation of genes v. UMI
metadata %>% 
  ggplot(aes(x=nUMI, y=nGene, color=mitoRatio)) + 
  geom_point() + 
  scale_colour_gradient(low = "gray90", high = "black") +
  stat_smooth(method=lm) +
  scale_x_log10() + 
  scale_y_log10() + 
  theme_classic() +
  geom_vline(xintercept = 750) +
  geom_hline(yintercept = 500) +
  facet_wrap(~sample)
```

## Filtering Cells Based on QC
```{r Filter}
#Subset high-quality cells on Determined QC metrics from above
filtered_seurat <- subset(x = merged_seurat, 
                          subset= (nUMI >= 750) & 
                            (nGene >= 500) & 
                            (log10GenesPerUMI > 0.80) & 
                            (mitoRatio < 0.10))

#Keep genes expressed in 10 or more cells
counts <- GetAssayData(object = filtered_seurat, slot = "counts")
nonzero <- counts > 0
keep_genes <- Matrix::rowSums(nonzero) >= 0
filtered_counts <- counts[keep_genes, ]

#Create Seurat Object after Filtering Low Quality Cells
filtered_seurat <- CreateSeuratObject(filtered_counts, meta.data = filtered_seurat@meta.data)
filtered_seurat
```
A total of 83 cells were removed. Our filtered dataset therefore has 7,648 cells with 55,536 unique features.  

## Determining Cell Cycle State
Cell cycle can be a confounding source of variability within our datasets. Therefore, we scored each cell based on expression of G2M and S phase features following the Cell-Cycle Scoring Vignette in Seurat. 

```{r Cell Cycle, warning=FALSE}
#Obtain list of mouse cell cycle genes
cc_file <- getURL("https://raw.githubusercontent.com/hbc/tinyatlas/master/cell_cycle/Mus_musculus.csv") 
cell_cycle_genes <- read.csv(text = cc_file)

#Determine Cell Cycle Score
Seurat_CellCycle <- NormalizeData(filtered_seurat)
Seurat_CellCycle <- CellCycleScoring(Seurat_CellCycle,
                                     g2m.features = cc.genes$g2m.genes,
                                     s.features = cc.genes$s.genes
)

#View(Seurat_CellCycle@meta.data)
Seurat_CellCycle <- FindVariableFeatures(Seurat_CellCycle, verbose = F)
Seurat_CellCycle <- ScaleData(Seurat_CellCycle, verbose = F)
Seurat_CellCycle <- RunPCA(Seurat_CellCycle, verbose = F)
DimPlot(Seurat_CellCycle, reduction = "pca", group.by = "Phase", split.by = "Phase")
```


As seen in the PCA plot, cell cycle does not appear to be a major source of variation in our dataset overall.

## SCTransform Normalization and Integration of Datasets
Normalization and variance stabilization of molecular count data was carried out using the R package sctransform (v0.2.1). We did not regress out mitochondrial genes, as this may mask important gene expression changes in efferocytosis. Since our lab has previously shown that macrophage metabolism shifts upon AC engulfment, we predict mitochondrial gene changes within our data.

```{r include=FALSE}
options(future.globals.maxSize= 993718400)
```


```{r SCTransform, message=FALSE, warning=FALSE}
#Split merged filtered Seurat into individual Seurat Objects
split_seurat <- SplitObject(filtered_seurat, split.by = "sample")
split_seurat <- split_seurat[c("WT_Ctrl", "WT_2_hr", "WT_6_hr", "MerKD_Ctrl", "MerKD_2hr", "MerKD_6_hr")]

#scTransform data without regressing out mitochondrial genes
for (i in 1:length(split_seurat)) {
  split_seurat[[i]] <- CellCycleScoring(split_seurat[[i]], g2m.features = cc.genes$g2m.genes, s.features = cc.genes$s.genes)
  split_seurat[[i]] <- SCTransform(split_seurat[[i]], verbose = F)
}

```

Integration of the datasets was carried out using IntegrateData function in Seurat. Importantly, the SCT-transformed datasets were prepared for integration using the PrepSCTIntegration function.

```{r Integration, echo=TRUE, warning=FALSE}
#Prepare and Integrate SCTransformed data into an integrated Seurat Object
#integ_features <- SelectIntegrationFeatures(object.list = split_seurat, nfeatures = 3000)
#split_seurat <- PrepSCTIntegration(object.list = split_seurat, anchor.features = integ_features, verbose = F)
#integ_anchors <- FindIntegrationAnchors(object.list = split_seurat, normalization.method = "SCT", anchor.features = integ_features, verbose = F)
#Integrated_Seurat <- IntegrateData(anchorset = integ_anchors, normalization.method = "SCT", verbose = F)
```

```{r include=FALSE}
Integrated_Seurat <- readRDS("results/Integrated_Seurat.rds")
```

```{r message=FALSE, warning=FALSE}
#Run PCA on integrated dataset
Integrated_Seurat <- RunPCA(object = Integrated_Seurat)
PCAPlot(Integrated_Seurat, split.by = "sample")
DimHeatmap(Integrated_Seurat, dims = 1:9, cells = 500, balanced = T)
print(x = Integrated_Seurat[["pca"]], dims = 1:10, nfeatures = 5)
ElbowPlot(object = Integrated_Seurat, ndims = 50)

#Run UMAP on integrated dataset
Integrated_Seurat <- RunUMAP(Integrated_Seurat, dims = 1:40)
DimPlot(Integrated_Seurat)

```


# Cluster Identification and Macrophage Heterogeneity at Steady State
## Seurat Clustering 
Clusters were identified using graph-based clustering approach implemented by the FindCluster function in Seurat at various resolutions to determine transcriptionally distinct populations of resident peritoneal cells. Clusters with high cell cycle scores were excluded from further downstream analyses. 

```{r}
Integrated_Seurat <- FindNeighbors(object = Integrated_Seurat, dims = 1:40, verbose = F)
Integrated_Seurat <- FindClusters(object = Integrated_Seurat, resolution = c(0.2, 0.4, 0.6, 0.8, 1.0, 1.4), verbose = F)

#Resolution 0.2
Idents(object = Integrated_Seurat) <- "integrated_snn_res.0.2"
Integrated_Seurat <- RunUMAP(Integrated_Seurat, reduction = "pca", dims = 1:40, verbose = F)
p7 <- DimPlot(Integrated_Seurat, reduction = "umap", label = T) +NoLegend()

#Resolution 0.4
Idents(object = Integrated_Seurat) <- "integrated_snn_res.0.4"
Integrated_Seurat <- RunUMAP(Integrated_Seurat, reduction = "pca", dims = 1:40, verbose = F)
p8 <- DimPlot(Integrated_Seurat, reduction = "umap", label = T, label.size = 6) +NoLegend()


#Resolution 0.6
Idents(object = Integrated_Seurat) <- "integrated_snn_res.0.6"
Integrated_Seurat <- RunUMAP(Integrated_Seurat, reduction = "pca", dims = 1:40, verbose = F)
p9 <- DimPlot(Integrated_Seurat, reduction = "umap", label = T, label.size = 6)  +NoLegend()


#Resolution 0.8
Idents(object = Integrated_Seurat) <- "integrated_snn_res.0.8"
Integrated_Seurat <- RunUMAP(Integrated_Seurat, reduction = "pca", dims = 1:40, verbose = F)
p10 <- DimPlot(Integrated_Seurat, reduction = "umap", label = T, label.size = 6) +NoLegend()

grid.arrange(p7,p8,p9,p10, ncol = 2)

#Revisit QC metrics
metrics <- c("nUMI", "nGene", "S.Score", "G2M.Score", "mitoRatio")
FeaturePlot(Integrated_Seurat, reduction = "umap", features = metrics, pt.size = 0.4, sort.cell = T, min.cutoff = "q10", label = T)

```

##SingleR Annotiations
To unbiasedly identify resident peritoneal cells present in the dataset, SingleR (v1.0.5) was employed. Briefly, SingleR infers the origin of each individual cell by referencing transcriptomic datasets of pure cell types. We utilized the ImmGen database, which contains normalized expression values for immune cells from 830 murine microarrays to ID our peritoneal cell types. 

```{r message=FALSE}
#Load SingleR Package
library(SingleR)
immgen.se <- ImmGenData()
```

```{r Single R, message=FALSE, warning=FALSE}
#Label Cells with Main Categories in ImmGen Database
counts <- GetAssayData(Integrated_Seurat)
pred.integrated <- SingleR(test = counts, ref = immgen.se, labels = immgen.se$label.main)
Main.labels <- table(pred.integrated$labels)
all.markers <- metadata(pred.integrated)$de.genes
```

```{r message=FALSE, warning=FALSE}
#Label Seurat object with SingleR annotations
to.remove <- pruneScores(pred.integrated)
summary(to.remove)

Integrated_Seurat[["SingleR.labels"]] <- pred.integrated$labels
DimPlot(Integrated_Seurat, group.by = "SingleR.labels", cols = c("royalblue", "forestgreen", "darkolivegreen3", "deepskyblue3", "gray", "navy", "orchid", "darkorchid4", "firebrick", "goldenrod", "azure4", "black", "darkgoldenrod4", "olivedrab1",  "tan2", "azure3", "forestgreen", "salmon"))
SingleR_markers <- metadata(pred.integrated)$de.genes

#Create Heatmap based on SingleR classifications
plotScoreHeatmap(pred.integrated)

```


## Macrophage Identification and Subsetting
### Macrophage Identification
To confirm the classification of macrophages in our dataset, we determined canonical immunne cell marker expression. 
```{r Immune Cell Marker Expression}
#Low resolution
Idents(object = Integrated_Seurat) <- "integrated_snn_res.0.2"
DimPlot(Integrated_Seurat, label = F)

DotPlot(Integrated_Seurat, features = c("Itgam", "Adgre1", "Icam2", "Itgax", "H2-Ab1", "Cd19", "Cd79a", "Cr2", "Cd3e", "Cd4", "Cd8a"))

FeaturePlot(Integrated_Seurat, features = c("Itgam", "Adgre1", "Icam2", "Itgax", "H2-Ab1", "Mertk"), order = T)
```
> Itgam (Cd11b), Itgax (Cd11c), Adgre1 (F4/80), H2-Ab1 (MHCII)

### Filtering Macrophages
Next we subsetted Clusters "0" and "3" as these were identified as macrophages and DCs by both SingleR and canonical macrophage marker expression. Additionally, we subset just macrophages in our MerTK +/+ Control condition to uncover peritoneal macrophage heterogenity at steady state.
```{r Subset Macs and WT Control Macs}
#Subset macrophages/DCs for all samples
Integrated_Seurat_MacsDCs <- SubsetData(Integrated_Seurat, ident.use = c("0", "3"))

#Subset Wild Type Macs only
Idents(Integrated_Seurat_MacsDCs) <- "sample"
WT_Control <- SubsetData(Integrated_Seurat_MacsDCs, ident.use = "WT_Ctrl")
Idents(WT_Control) <- "integrated_snn_res.0.4"
DimPlot(WT_Control, label = T)
WT_Control_Macs <- SubsetData(WT_Control, ident.use = c("1", "3", "5", "6", "11"))
```

Note that Cluster "12" was omitted from downstream analyses as this cluster was enriched for cell G2M cell cycle genes compared to other macrophage clusters.

### UMAP Plot of Macrophage Clusters
```{r include=FALSE}
WT_Control_vis <- readRDS("results/WT_Control_vis.rds")
```

```{r DimPlot in Figure 1B}
#Remove outlier cells for visualization
#Macsplot <- DimPlot(WT_Control)
#Macs_selection <- CellSelector(Macsplot)
#Idents(WT_Control, cells = Macs_selection) <- "Macrophages"
#WT_Control_vis <- SubsetData(WT_Control, ident.use = "Macrophages")

#Dimplot in Figure 1B
Idents(object = WT_Control_vis) <- "integrated_snn_res.0.4"
DimPlot(WT_Control_vis, cols = c("royalblue3", "firebrick", "deepskyblue2", "goldenrod", "darkorchid", "gray"), pt.size = 2)
```

### Heatmap of Macrophage Clusters
```{r Heatmap of Macrophage Clusters Figure 1C, message=FALSE, warning=FALSE}
#Determine Top 50 markers for each macrophage cluster
DefaultAssay(WT_Control_Macs) <- "RNA"
WT_Control_Macs  <- NormalizeData(WT_Control_Macs, verbose = F)

WT_Control_Macs_markers <- FindAllMarkers(WT_Control_Macs , only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, test.use = "MAST", verbose = F)

WT_Control_Macs_markers_top50 <- WT_Control_Macs_markers %>% group_by(cluster) %>% top_n(n = 50, wt = avg_logFC)
head(WT_Control_Macs_markers_top50)

#Heatmap in Figure 1C
DoHeatmap(WT_Control_Macs, features = WT_Control_Macs_markers_top50$gene, slot = "scale.data", assay = "SCT", group.colors = c("royalblue3", "firebrick", "deepskyblue2", "goldenrod", "darkorchid"),  disp.min = -2) + NoLegend()
```

### Plotting Features in Macrophage UMAP
```{r Feature Plot Mac Markers}
DefaultAssay(WT_Control_vis) <- "RNA"
FeaturePlot(WT_Control_vis, features = c("Itgam", "Adgre1", "Icam2", "Timd4", "Klf2", "Mertk", "Itgax", "H2-Ab1", "Ccr2"), order = T, pt.size = 2, ncol = 3)
```

### Plotting Features in Macrophage Violin Plot
```{r Violin Plot Mac Markers}
#Subset Macrophages further in WT Steady State Condition
Idents(WT_Control) <- "integrated_snn_res.0.4"
WT_Control_Macs <- SubsetData(WT_Control, ident.use = c("1", "3", "5", "6"))

#Cytokine and Marker Gene Expression - VlnPlot in Figure 1E
DefaultAssay(WT_Control_Macs) <- "SCT"
VlnPlot.Macs <- VlnPlot(WT_Control_Macs, features = c("Klf2", "Gata6", "Cd93", "Csf1r", "Anxa5", "Cd36", "Actb", "Arg1", "Irf4", "Retnla", "Cd74", "Tgfbi",  "Il1b", "Il6", "Tnf", "Cxcl2"),  cols = c("royalblue3", "firebrick", "deepskyblue2", "goldenrod"), pt.size = 0, combine = F)
for(i in 1:length(VlnPlot.Macs)) {
  VlnPlot.Macs[[i]] <- VlnPlot.Macs[[i]] + NoLegend() + theme(axis.title.y = element_blank(), axis.title.x = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank())
}
cowplot::plot_grid(plotlist = VlnPlot.Macs, ncol = 4)

```








### Differential Gene Expression Between Clusters
Differentially expressed genes were determined using the MAST algorithm between Cluster 1 and other macrophage clusters. Once determined, the differnetially expressed genes were submitted to gProfiler for gene ontology.

```{r Cluster Marker Comaprison, message=FALSE}
#Cluster Comparisons for gProfiler###############################
Mac1_3 <- FindMarkers(WT_Control_Macs, ident.1 = c("1"), ident.2 = "3", test.use = "MAST", min.pct = 0.25, logfc.threshold = 0.25, verbose = F)
Mac1_3.Markers <- subset(Mac1_3, Mac1_3$p_val_adj < 0.05)
head(Mac1_3.Markers)

Mac1_5 <- FindMarkers(WT_Control_Macs, ident.1 = c("1"), ident.2 = "5", test.use = "MAST", min.pct = 0.25, logfc.threshold = 0.25, verbose = F)
Mac1_5.Markers <- subset(Mac1_5, Mac1_5$p_val_adj < 0.05)
head(Mac1_5.Markers)

Mac1_6 <- FindMarkers(WT_Control_Macs, ident.1 = "1", ident.2 = "6", test.use = "MAST", min.pct = 0.25, logfc.threshold = 0.25, verbose = F)
Mac1_6.Markers <- subset(Mac1_6, Mac1_6$p_val_adj < 0.05)
head(Mac1_6.Markers)
```

# Efferocytosis Transcriptional Signatures
Peritoneal macrophages were cultured with apoptotic cells in an ex vivo efferocytosis assay. Apoptotic cells (ACs) were generated using GFP-Jurkat Human T cells (GenTarget) exposed to UV radiation for 7 minutes followed by a 2-hour incubation at 37°C. Apoptosis was confirmed by annexin V positive, propidium iodide negative identification. Adherent resident peritoneal cells were co-cultivated with ACs at a ratio of 5 ACs to 1 peritoneal cell for 2 hours or 6 hours as indicated. Control cells were given a media change corresponding to the 6-hour timepoint. Non-engulfed ACs were removed from the co-culture through rigorous rinsing with warm saline. 

```{r Subset Data for Efferocytosis}
#Isolate Wild Type Macrophages from all time points
Idents(Integrated_Seurat) <- "integrated_snn_res.0.4"
Integrated_Macs <- SubsetData(Integrated_Seurat, ident.use = c("1", "3", "5", "6"))
```
```{r include=FALSE}
Integrated_Macs <- readRDS("results/Integrated_Macs_vis")
```
```{r}
#Remove "outlier" cells
#Integrated_Macsplot <- DimPlot(Integrated_Macs)
#Integrated_Macs_selection <- CellSelector(Integrated_Macsplot)
#Idents(Integrated_Macs, cells = Integrated_Macs_selection) <- "Macrophages"
#Integrated_Macs <- SubsetData(Integrated_Macs, ident.use = "Macrophages")

Idents(Integrated_Macs) <- "integrated_snn_res.0.4"
DimPlot(Integrated_Macs)

#Isolate Wild Type F4/80+ Peritoneal Macrophages
Idents(Integrated_Macs) <- "sample"
Integrated_Macs_WT <- SubsetData(Integrated_Macs, ident.use = c("WT_Ctrl", "WT_2_hr", "WT_6_hr"))

#Control
p11 <- DimPlot(Integrated_Macs_WT, group.by = "sample", cols = c("grey", "grey", "blue"), pt.size = 3, order = T)

#6 hour
p12 <- DimPlot(Integrated_Macs_WT, group.by = "sample", cols = c("grey", "blue", "grey"), pt.size = 3, order = F)

#2 hour
Integrated_Macs_WT$sample <- factor(Integrated_Macs_WT$sample, levels = c("WT_Ctrl", "WT_6_hr", "WT_2_hr"))
p13 <- DimPlot(Integrated_Macs_WT, group.by = "sample", cols = c("grey", "grey", "blue"), pt.size = 3, order = T)

grid.arrange(p11, p12, p13, ncol = 1)
```

## Pseudotime Trajectory Analysis
With our inference that each cell represented in our analysis was at a unique stage of efferocytosis, we performed a pseudotemporal reconstruction using Slingshot, a tool that infers global lineage structures and pseudotime variables for each cell along a lineage. Two distinct lineages were observed which corresponded to unique transcriptional changes in our F4/80+ LPMs and MHCII+ SPMs. 

### Pseudotime with Slingshot
```{r Pseudotime, message=FALSE, warning=FALSE}
#Isolate Macrophage Subpopulations
Idents(Integrated_Macs_WT) <- "integrated_snn_res.0.4"
Integrated_Macs_WT_MHCIILO <- SubsetData(Integrated_Macs_WT, ident.use = c("1", "3", "5"))
Integrated_Macs_WT_MHCIIHI <- SubsetData(Integrated_Macs_WT, ident.use = "6")

#Create single cell experiment objects
Integrated_Macs_WT_MHCIILO_SCE <- as.SingleCellExperiment(Integrated_Macs_WT_MHCIILO)
table(Integrated_Macs_WT_MHCIILO_SCE$sample)

Integrated_Macs_WT_MHCIIHI_SCE <- as.SingleCellExperiment(Integrated_Macs_WT_MHCIIHI)
table(Integrated_Macs_WT_MHCIIHI_SCE$sample)

#Re-order sample levels
Integrated_Macs_WT_MHCIILO_SCE$sample <- factor(Integrated_Macs_WT_MHCIILO_SCE$sample, levels = c("WT_2_hr", "WT_6_hr", "WT_Ctrl"))

Integrated_Macs_WT_MHCIIHI_SCE$sample <- factor(Integrated_Macs_WT_MHCIIHI_SCE$sample, levels = c("WT_2_hr", "WT_6_hr", "WT_Ctrl"))

#Assign Cluster IDs in SCE object
Idents(Integrated_Macs_WT_MHCIILO) <- "integrated_snn_res.0.4"
colData(Integrated_Macs_WT_MHCIILO_SCE)$Seurat_clusters <- as.character(Integrated_Macs_WT_MHCIILO@active.ident)

Idents(Integrated_Macs_WT_MHCIIHI) <- "integrated_snn_res.0.4"
colData(Integrated_Macs_WT_MHCIIHI_SCE)$Seurat_clusters <- as.character(Integrated_Macs_WT_MHCIIHI@active.ident)

###Run Slingshot Analysis on Different Populations of WT Macs
#MHCII LO UMAP Trajectory
Integrated_Macs_WT_MHCIILO_SCE <- slingshot(Integrated_Macs_WT_MHCIILO_SCE, clusterLabels = "Seurat_clusters", reducedDim = "UMAP")
colors <- rainbow(100, alpha = 1)
plot(reducedDims(Integrated_Macs_WT_MHCIILO_SCE)$UMAP, col = colors[cut(Integrated_Macs_WT_MHCIILO_SCE$slingPseudotime_1, breaks=100)], pch=16, asp = 1, cex = 0.5, main = "MHCIILO Macrophages")
lines(SlingshotDataSet(Integrated_Macs_WT_MHCIILO_SCE), lwd=4, col ="black")

#MHCII HI UMAP Trajectory
Integrated_Macs_WT_MHCIIHI_SCE <- slingshot(Integrated_Macs_WT_MHCIIHI_SCE, clusterLabels = "Seurat_clusters", reducedDim = "UMAP")
colors <- rainbow(100, alpha = 1)
plot(reducedDims(Integrated_Macs_WT_MHCIIHI_SCE)$UMAP, col = colors[cut(Integrated_Macs_WT_MHCIIHI_SCE$slingPseudotime_1, breaks=100)], pch=16, asp = 1, cex = 0.5, main = "MHCIIHI Macrophages")
lines(SlingshotDataSet(Integrated_Macs_WT_MHCIIHI_SCE), lwd=4, col ="black")
```


Next, we ran slingshot with PCA as the method for reduced dimension as UMAP can incorrectly over-reduce dimension potentially obscurring data from pseudotime. 

```{r PCA Pseudotime, message=FALSE}
#PCA MHCII LO
Integrated_Macs_WT_MHCIILO_SCE <- slingshot(Integrated_Macs_WT_MHCIILO_SCE, clusterLabels = "Seurat_clusters", reducedDim = "PCA")

colors <- rainbow(100, alpha = 1)
plot(reducedDims(Integrated_Macs_WT_MHCIILO_SCE)$PCA, col = colors[cut(Integrated_Macs_WT_MHCIILO_SCE$slingPseudotime_1, breaks=100)], pch=16, asp = 1, cex = 0.5, main = "MHCIILO Macrophages")
lines(SlingshotDataSet(Integrated_Macs_WT_MHCIILO_SCE), lwd=4, col ="black")

ggplot(as.data.frame(colData(Integrated_Macs_WT_MHCIILO_SCE)), aes(x = Integrated_Macs_WT_MHCIILO_SCE$slingPseudotime_1, y = sample, colour = Seurat_clusters)) +  geom_quasirandom(groupOnX = FALSE) +  scale_color_tableau() + theme_classic() +  xlab("Slingshot pseudotime") + ylab("Timepoint") + ggtitle("MHCIILO Macs by Condition")


#PCA MHCII HI
Integrated_Macs_WT_MHCIIHI_SCE <- slingshot(Integrated_Macs_WT_MHCIIHI_SCE, clusterLabels = "Seurat_clusters", reducedDim = "PCA")

colors <- rainbow(100, alpha = 1)
plot(reducedDims(Integrated_Macs_WT_MHCIIHI_SCE)$PCA, col = colors[cut(Integrated_Macs_WT_MHCIIHI_SCE$slingPseudotime_1, breaks=100)], pch=16, asp = 1, cex = 0.5, main = "MHCIIHI Macrophages")
lines(SlingshotDataSet(Integrated_Macs_WT_MHCIIHI_SCE), lwd=4, col ="black")

ggplot(as.data.frame(colData(Integrated_Macs_WT_MHCIIHI_SCE)), aes(x = Integrated_Macs_WT_MHCIIHI_SCE$slingPseudotime_1, y = sample, colour = sample)) +  geom_quasirandom(groupOnX = FALSE) +  scale_color_tableau() + theme_classic() +  xlab("Slingshot pseudotime") + ylab("Timepoint") + ggtitle("MHCIIHI Macs by Condition")

```

### Determining Differnetially Expressed Genes After Efferocytosis
```{r Effero DE Genes, message=FALSE}
###DE Genes
#MHCII LO
DefaultAssay(Integrated_Macs_WT_MHCIILO) <- "RNA"
Integrated_Macs_WT_MHCIILO <- NormalizeData(Integrated_Macs_WT_MHCIILO)
Idents(Integrated_Macs_WT_MHCIILO) <- "sample"
WT_MHCIILO_Markers <- FindMarkers(Integrated_Macs_WT_MHCIILO, ident.1 =  "WT_6_hr", ident.2 = "WT_Ctrl", test.use = "MAST")
WT_MHCIILO_Markers <- subset(WT_MHCIILO_Markers, WT_MHCIILO_Markers$p_val < 0.05 & abs(WT_MHCIILO_Markers$avg_logFC) > 0.4)

#MHCII HI
DefaultAssay(Integrated_Macs_WT_MHCIIHI) <- "RNA"
Integrated_Macs_WT_MHCIIHI <- NormalizeData(Integrated_Macs_WT_MHCIIHI)
Idents(Integrated_Macs_WT_MHCIIHI) <- "sample"
WT_MHCIIHI_Markers <- FindMarkers(Integrated_Macs_WT_MHCIIHI, ident.1 =  "WT_6_hr", ident.2 = "WT_Ctrl", test.use = "MAST")
WT_MHCIIHI_Markers <- subset(WT_MHCIIHI_Markers, WT_MHCIIHI_Markers$p_val < 0.05 & abs(WT_MHCIIHI_Markers$avg_logFC) > 0.4)

#Plot Expression of Interesting DEgenes as a function of pseudotime
plotExpression(Integrated_Macs_WT_MHCIILO_SCE, features = c("Cd36", "Il1b", "Il6", "Tgfbi", "H2-Ab1", "Il1rn"), x = "slingPseudotime_1", show_smooth = T, point_alpha = 0.4, colour_by = "sample", scales = "free", show_se = F) 

plotExpression(Integrated_Macs_WT_MHCIIHI_SCE, features = c("Cd36", "Il1b", "Il6", "Tgfbi", "H2-Ab1", "Il1rn"), x = "slingPseudotime_1", show_smooth = T, colour_by = "sample", scales = "free", show_se = F) 
```

### Creating Dataset to import into Phantasus
```{r CLuster Averages Effero, message=FALSE}
#MHCII LO
Idents(Integrated_Macs_WT_MHCIILO) <- "sample"
cluster.averages <- AverageExpression(Integrated_Macs_WT_MHCIILO)
write.csv(cluster.averages[["SCT"]], file = "./results/Integrated_Macs_WT_MHCIILO.csv", row.names = TRUE, quote = FALSE)

#MHCII HI
Idents(Integrated_Macs_WT_MHCIIHI) <- "sample"
cluster.averages <- AverageExpression(Integrated_Macs_WT_MHCIIHI)
write.csv(cluster.averages[["SCT"]], file = "./results/Integrated_Macs_WT_MHCIIHI.csv", row.names = TRUE, quote = FALSE)
```

### Gene Ontolgy using GOPlot
```{r Gene Ontology MHCII LO}
#MHCII LO
## Organize Selected Gene Pathways from gProfiler into a compatible data frame
WT_MHCII_LO_GO_pathways <- read.csv(file = "./results/gProfiler Results/WT_MHCIILO.csv")
WT_MHCII_LO_GO_pathways <- as.data.frame(WT_MHCII_LO_GO_pathways)

## Organize Differnetially Expressed genes into compatible data frame
WT_MHCIILO_Markers_genes <- as.data.frame(WT_MHCIILO_Markers)
WT_MHCIILO_Markers_genes$ID <- row.names(WT_MHCIILO_Markers_genes)
names(WT_MHCIILO_Markers_genes)[1] <- "P.Value"
names(WT_MHCIILO_Markers_genes)[2] <- "logFC"
names(WT_MHCIILO_Markers_genes)[5] <- "adj.P.Val"
head(WT_MHCIILO_Markers_genes)

##Combine datasets above to form an object for GOPlot
WT_MHCII_LO_circ <- circle_dat(WT_MHCII_LO_GO_pathways, WT_MHCIILO_Markers_genes)

##Create a Bubble Plot
GOBubble(WT_MHCII_LO_circ, ID = F, table.legend = F, colour = c("gray25", "gray25"), labels = 2)
```

```{r Gene Ontology MHCII HI}
#MHCII HI
## Organize Selected Gene Pathways from gProfiler into a compatible data frame
WT_MHCII_HI_GO_pathways <- read.csv(file = "./results/gProfiler Results/WT_MHCIIHI.csv")
WT_MHCII_HI_GO_pathways <- as.data.frame(WT_MHCII_HI_GO_pathways)

## Organize Differnetially Expressed genes into compatible data frame
WT_MHCIIHI_Markers_genes <- as.data.frame(WT_MHCIIHI_Markers)
WT_MHCIIHI_Markers_genes$ID <- row.names(WT_MHCIIHI_Markers_genes)
names(WT_MHCIIHI_Markers_genes)[1] <- "P.Value"
names(WT_MHCIIHI_Markers_genes)[2] <- "logFC"
names(WT_MHCIIHI_Markers_genes)[5] <- "adj.P.Val"
head(WT_MHCIIHI_Markers_genes)

##Combine datasets above to form an object for GOPlot
WT_MHCII_HI_circ <- circle_dat(WT_MHCII_HI_GO_pathways, WT_MHCIIHI_Markers_genes)

##Create a Bubble Plot
GOBubble(WT_MHCII_HI_circ, ID = F, table.legend = F, colour = c("goldenrod", "goldenrod"), labels = 2)
```


# MerTK KO Steady State
To validate our approach, we incorporated a well characterized efferocytosis receptor knockout, Mertk-/-. We performed the same experimental design as above.

## Determine Differences at Steady State between MerTK +/+ and MerTK -/- Macs
```{r Subset MerTK KO Macs, message=FALSE}
#Isolate MerTK -/- (MerKD) Macrophages
Idents(Integrated_Macs) <- "sample"
Integrated_Macrophages_MerKD <- SubsetData(Integrated_Macs, ident.use = c("MerKD_Ctrl", "MerKD_2hr", "MerKD_6_hr"))

#Isolate Steady State Macs of MerTK +/+ and MerTK -/- 
Integrated_Macrophages_Control <- SubsetData(Integrated_Macs, ident.use = c("WT_Ctrl", "MerKD_Ctrl"))

#UMAP Plot at Steady State
Idents(Integrated_Macrophages_Control) <- "integrated_snn_res.0.4"
DimPlot(Integrated_Macrophages_Control, split.by = "sample", cols = c("royalblue3", "firebrick", "deepskyblue2", "goldenrod", "darkorchid"))
```

### Differential Gene Expression in MerTK -/- Macs
```{r DE Genes MerTK KO Macs, message=FALSE}
DefaultAssay(Integrated_Macrophages_Control) <- "RNA"
Integrated_Macrophages_Control  <- NormalizeData(Integrated_Macrophages_Control)
Idents(Integrated_Macrophages_Control) <- "sample"
Control_Markers <- FindMarkers(Integrated_Macrophages_Control, ident.1 = "WT_Ctrl", ident.2 = "MerKD_Ctrl", test.use = "MAST")
Control_Markers_pos <- subset(Control_Markers, Control_Markers$p_val_adj < 0.05 & Control_Markers$avg_logFC > 0)
Control_Markers_neg <- subset(Control_Markers, Control_Markers$p_val_adj < 0.05 & Control_Markers$avg_logFC < 0)
head(Control_Markers_pos, n = 10)
head(Control_Markers_neg, n =10)
```

### Determine how MerTK KO affects specific Macrophage Subsets
Since MerTK is predominately expressed on F4/80+ MHCII Lo Macrophages, we determined how MerTK KO affected expression of cytokines and other macrophage markers at steady state.

```{r Group Macs Together}
Idents(Integrated_Macrophages_Control) <- "integrated_snn_res.0.4"
Integrated_Macrophages_Control_Subset <- SubsetData(Integrated_Macrophages_Control, ident.use = c("1", "3", "5", "6"))

#Rename cluster IDs to represent either F4/80+ or MHCII HI
Integrated_Macrophages_Control_Subset <- RenameIdents(Integrated_Macrophages_Control_Subset, "1" =  "F4/80+", "3" = "F4/80+", "5" = "F4/80+", "6" = "MHCII+")
Integrated_Macrophages_Control_Subset$sample <- factor(Integrated_Macrophages_Control_Subset$sample, levels = c("WT_Ctrl", "MerKD_Ctrl"))

#Plot changes in Selected Genes
VlnPlot.Macs <- VlnPlot(Integrated_Macrophages_Control_Subset, features = c("Il1b", "Tgfbi", "Axl", "H2-Ab1", "Cd74", "Cd36"), pt.size = 0, cols = c("firebrick", "grey"), combine = F, split.by = "sample", sort = "decreasing")
for(i in 1:length(VlnPlot.Macs)) {
  VlnPlot.Macs[[i]] <- VlnPlot.Macs[[i]] + NoLegend()+ theme(axis.title.y = element_blank(), axis.title.x = element_blank(), axis.text.x = element_blank(), axis.text.y = element_blank(), axis.ticks.y = element_blank())
}
cowplot::plot_grid(plotlist = VlnPlot.Macs, ncol = 3)

#Feature Plot of Mac Markers in Figure 1 in MerTK KO
Idents(Integrated_Macrophages_Control) <- "sample"
Integrated_Macrophages_Control_MerKD <- SubsetData(Integrated_Macrophages_Control, ident.use = c("MerKD_Ctrl"))
FeaturePlot(Integrated_Macrophages_Control_MerKD, features = c("Itgam", "Adgre1", "Icam2", "Itgax", "H2-Ab1", "Ccr2", "Timd4", "Klf2", "Mertk"), order = T, pt.size = 1.5, ncol = 3)
```

## Efferocytosis in MerTK -/-
### Experimental Time Points
```{r Effero Time Points}
Idents(Integrated_Macrophages_MerKD) <- "sample"
#6 Hour 
Integrated_Macrophages_MerKD$sample <- factor(Integrated_Macrophages_MerKD$sample, levels = c("MerKD_Ctrl", "MerKD_2hr", "MerKD_6_hr"))
DimPlot(Integrated_Macrophages_MerKD, group.by = "sample", cols = c("grey", "grey", "firebrick3"), pt.size = 3, order = T)
#2 Hour
Integrated_Macrophages_MerKD$sample <- factor(Integrated_Macrophages_MerKD$sample, levels = c("MerKD_Ctrl", "MerKD_6_hr", "MerKD_2hr"))
DimPlot(Integrated_Macrophages_MerKD, group.by = "sample", cols = c("grey", "grey", "firebrick3"), pt.size = 3, order = T)
#Control
Integrated_Macrophages_MerKD$sample <- factor(Integrated_Macrophages_MerKD$sample, levels = c("MerKD_2hr", "MerKD_6_hr", "MerKD_Ctrl"))
DimPlot(Integrated_Macrophages_MerKD, group.by = "sample", cols = c("grey", "grey", "firebrick3"), pt.size = 3, order = T)
```
The F4/80+ macrophages appear to be significantly altered after challenge with ACs. There we looked at this population further to determine how Mertk KO altered the transcriptional program of this macrophage subset during efferocytosis.

### Slingshot Trajectory Analyses
```{r}
#Subset F4/80+ Macs
Idents(Integrated_Macrophages_MerKD) <- "integrated_snn_res.0.4"
Integrated_Macrophages_MerKD_F480 <- SubsetData(Integrated_Macrophages_MerKD, ident.use = c("1", "3", "5"))

#Create single cell experiment object
Integrated_Macrophages_MerKD_F480_SCE <- as.SingleCellExperiment(Integrated_Macrophages_MerKD_F480)
table(Integrated_Macrophages_MerKD_F480_SCE$sample)

#Re-order sample levels
Integrated_Macrophages_MerKD_F480_SCE$sample <- factor(Integrated_Macrophages_MerKD_F480_SCE$sample, levels = c("MerKD_6_hr", "MerKD_2hr", "MerKD_Ctrl"))

#Assign Cluster IDs in SCE object
Idents(Integrated_Macrophages_MerKD_F480) <- "integrated_snn_res.0.4"
colData(Integrated_Macrophages_MerKD_F480_SCE)$Seurat_clusters <- as.character(Integrated_Macrophages_MerKD_F480@active.ident)

###Run Slingshot Analysis
#UMAP
Integrated_Macrophages_MerKD_F480_SCE <- slingshot(Integrated_Macrophages_MerKD_F480_SCE, clusterLabels = "Seurat_clusters", reducedDim = "UMAP")

colors <- rainbow(100, alpha = 1)
plot(reducedDims(Integrated_Macrophages_MerKD_F480_SCE)$UMAP, col = colors[cut(Integrated_Macrophages_MerKD_F480_SCE$slingPseudotime_1, breaks=100)], pch=16, asp = 1, cex = 0.5, main = "F480 Macrophages")
lines(SlingshotDataSet(Integrated_Macrophages_MerKD_F480_SCE), lwd=4, col ="black", type = "lineages")

#PCA
Integrated_Macrophages_MerKD_F480_SCE <- slingshot(Integrated_Macrophages_MerKD_F480_SCE, clusterLabels = "Seurat_clusters", reducedDim = "PCA", start.clus = "1")

colors <- rainbow(50, alpha = 1)
plot(reducedDims(Integrated_Macrophages_MerKD_F480_SCE)$PCA, col = colors[cut(Integrated_Macrophages_MerKD_F480_SCE$slingPseudotime_1,breaks=50)], pch=16, asp = 1)
lines(SlingshotDataSet(Integrated_Macrophages_MerKD_F480_SCE), lwd=4, col ="black", type ="lineages")

#Pseudotime as a function of Cluster
ggplot(as.data.frame(colData(Integrated_Macrophages_MerKD_F480_SCE)), aes(x = Integrated_Macrophages_MerKD_F480_SCE$slingPseudotime_1, y = sample, colour = Seurat_clusters)) +  geom_quasirandom(groupOnX = FALSE) +  scale_color_tableau() + theme_classic() +  xlab("Slingshot pseudotime") + ylab("Timepoint") + ggtitle("F480 Macs by Condition") 
```

### Differential Gene Expression & Gene Ontology
```{r}
#DE Genes before and after 6 hour of Effero
DefaultAssay(Integrated_Macrophages_MerKD_F480) <- "RNA"
Integrated_Macrophages_MerKD_F480 <- NormalizeData(Integrated_Macrophages_MerKD_F480)

Idents(Integrated_Macrophages_MerKD_F480) <- "sample"
MerKD_F480_Markers <- FindMarkers(Integrated_Macrophages_MerKD_F480, ident.2 =  "MerKD_Ctrl", ident.1 =  "MerKD_6_hr", test.use = "MAST", only.pos = T)
MerKD_F480_Markers <- subset(MerKD_F480_Markers, MerKD_F480_Markers$p_val_adj < 0.05)
head(MerKD_F480_Markers, n = 10)

#WT MHCII LO
DefaultAssay(Integrated_Macs_WT_MHCIILO) <- "RNA"
Integrated_Macs_WT_MHCIILO <- NormalizeData(Integrated_Macs_WT_MHCIILO)

Idents(Integrated_Macs_WT_MHCIILO) <- "sample"
WT_MHCIILO_Markers <- FindMarkers(Integrated_Macs_WT_MHCIILO, ident.2 =  "WT_Ctrl", ident.1 = "WT_6_hr", test.use = "MAST", only.pos = T)
WT_MHCIILO_Markers <- subset(WT_MHCIILO_Markers, WT_MHCIILO_Markers$p_val < 0.05)
head(WT_MHCIILO_Markers)

#F480 Mac Comparison between genotypes
F480_Same_DEgenes <- intersect(row.names(MerKD_F480_Markers), row.names(WT_MHCIILO_Markers))
F480_Control_unique <- setdiff(row.names(WT_MHCIILO_Markers), row.names(MerKD_F480_Markers))
F480_Knockout_unique <- setdiff(row.names(MerKD_F480_Markers), row.names(WT_MHCIILO_Markers))
```

Gene Profiling was carried out on genes that were uniquely upregulated by each genotype along with genes that were upregulated by both genotypes.

```{r GO Plots MerTK KO}
#gProfiler Results
MHCIILO_Common_GO <- read.csv(file = "./results/gProfiler Results/MHCIILO_Common.csv")
MHCIILO_Common_GO <- as.data.frame(MHCIILO_Common_GO)
MerKD_MHCIILO_Unique_GO <- read.csv(file = "./results/gProfiler Results/MerKD_MHCIILO_Unique.csv")
MerKD_MHCIILO_Unique_GO <- as.data.frame(MerKD_MHCIILO_Unique_GO)
WT_MHCIILO_Unique_GO <- read.csv(file = "./results/gProfiler Results/WT_MHCIILO_Unique.csv")
WT_MHCIILO_Unique_GO <- as.data.frame(WT_MHCIILO_Unique_GO)


MerKD_F480_Markers_genes <- as.data.frame(MerKD_F480_Markers)
MerKD_F480_Markers_genes$ID <- row.names(MerKD_F480_Markers_genes)
names(MerKD_F480_Markers_genes)[1] <- "P.Value"
names(MerKD_F480_Markers_genes)[2] <- "logFC"
names(MerKD_F480_Markers_genes)[5] <- "adj.P.Val"
head(MerKD_F480_Markers_genes)

WT_MHCIILO_Markers_genes <- as.data.frame(WT_MHCIILO_Markers)
WT_MHCIILO_Markers_genes$ID <- row.names(WT_MHCIILO_Markers_genes)
names(WT_MHCIILO_Markers_genes)[1] <- "P.Value"
names(WT_MHCIILO_Markers_genes)[2] <- "logFC"
names(WT_MHCIILO_Markers_genes)[5] <- "adj.P.Val"
head(WT_MHCIILO_Markers_genes)

#Common
MHCIILO_Common_GO_circle <- circle_dat(MHCIILO_Common_GO, WT_MHCIILO_Markers_genes)
GOBubble(MHCIILO_Common_GO_circle , ID = F, table.legend = F, colour = c("gray40", "gray40"), labels = 0)

#MerKD Unique
MerKD_MHCIILO_Unique_GO_circle <- circle_dat(MerKD_MHCIILO_Unique_GO, MerKD_F480_Markers_genes)
GOBubble(MerKD_MHCIILO_Unique_GO_circle , ID = F, table.legend = F, colour = c("firebrick", "firebrick"), labels = 1)

#WT Unique
WT_MHCIILO_Unique_GO_circle <- circle_dat(WT_MHCIILO_Unique_GO, WT_MHCIILO_Markers_genes)
GOBubble(WT_MHCIILO_Unique_GO_circle , ID = F, table.legend = F, colour = c("gray25", "gray25"), labels = 0)
```
